Demo arXiv template

### 2022-11-03 paper/paper.qmd
change the small example to compare between two distributions

*first sentence: tailor to their indices without thinking about the big picture*

~~While every jumbled index jumbles in its own way, [perspicuous] clearly-laied out indices are all alike.~~
This paper proposes a data pipeline for index construction. 
By recognising the common steps shared by many indices, we develop a pipeline that breaks down index construction into multiple modules and allow operations in various modules to be combined like building blocks to construct indices.
The pipeline approach is general while adaptable to most index construction. 
It allows indices to be created, studied, and compared in a structured tidy form and enables statistical analysis of indices to be performed easily: 
More specifically, it enables researchers to 1) validate the indices calculated from external organisations, 2) unify various indices under the same framework for computing, 3) swap or adjust individual steps in the index construction to study their contribution, 4) calculate uncertainty on indices through bootstrap or others, 5) enhance existing indices through comparing and studying their statistical properties, and finally, 6) propose new indices from combining different steps in existing indices.

@@

# Data pipeline in R {#sec-data-pipeline-in-r}

@@

Consider a commonly used drought index: Standardised Precipitation Index (SPI) [@spi]. Its construction involves: 1) aggregating monthly precipitation (`prcp`) with a time scale, 2) fitting the aggregated series with a chosen distribution to get density values, and 3) converting the density values to normal quantiles. Under the pipeline approach, SPI will be constructed as:

```{r eval = FALSE, echo = TRUE}
DATA %>% 
  aggregate(scale = 12, var = prcp) %>% 
  normalise(dist = gamma, var = prcp, fit_method = lmom) %>% 
  augment(var = .agg_prcp)
```

@@

Here each command corresponds to one step described above with additional arguments specific to the step. 
The pipeline construction produces the same index value as command like `spi(...)`, as shown in @fig-toy-example, but with additional benefit:

@@

```{r eval = FALSE, echo = FALSE}
res <- w %>% 
  ts_aggregate(scale = 12, date = ym, id = id, col = prcp) %>% 
  normalising(dist = "gamma", date = ym, id = id, col = prcp,
              method = "lmom") %>% 
  augment(id = id, date = ym,  col = "prcp_agg")

res <- res %>% 
  rename(
         pipeline_d = prcp_agg,
         pipeline_idx = .index)

data(wichita)
wichita <- ts(wichita[,-c(1,2)], end=c(2011,10), frequency=12)
spi12 <- spi(wichita[,'PRCP'], 12)
from_SPI <- tibble(
  ym = yearmonth(zoo::yearmon(zoo::index(wichita))), 
  SPI_idx = as.vector(spi12$fitted)
)

compare <- res %>% left_join(from_SPI, by = "ym") 
compare %>% 
  ggplot(aes(x = SPI_idx, y = pipeline_idx)) + 
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") + 
  geom_point() + 
  theme_bw() + 
  theme(legend.position = "bottom", 
        aspect.ratio = 1) +
  xlab("Index values from `spi()` in package `SPEI`") + 
  ylab("Index values from pipeline") + 
  theme(panel.grid.minor = element_blank())
```

```{r fig-toy-example, eval = TRUE, echo = FALSE}
#| fig-align: center
#| fig-cap: "Standardised Precipitation Index (SPI) calculated using two approaches: the pipeline construction (y-axis) and the function `spi()` from the `SPEI` package (x-axis), using the `wichita` data from the package `SPEI`. All the points align diagnally, indicating the same numerical result from both approaches."
#| out-height: 70%
#| out-width: 70%
knitr::include_graphics(here::here("figures/toy-example-spi.png"))
```

@@

*put a diagram here to summarise the pipeline*

@@

each step as a point rather than subsection

>>>>

This paper proposes a data pipeline for index construction. 
By recognising the common steps shared by many indices, we develop a pipeline that breaks down index construction into multiple modules and allow operations in various modules to be combined like building blocks to construct indices.
The pipeline approach is general while adaptable to most index construction. 
It allows indices to be created, studied, and compared in a structured tidy form and enables statistical analysis of indices to be performed easily: 
More specifically, it enables researchers to 1) validate the indices calculated from external organisations, 2) unify various indices under the same framework for computing, 3) swap or adjust individual steps in the index construction to study their contribution, 4) calculate uncertainty on indices through bootstrap or others, 5) enhance existing indices through comparing and studying their statistical properties, and finally, 6) propose new indices from combining different steps in existing indices.

@@

# Data pipeline {#sec-data-pipeline-in-r}

*Think about if there is another word for data pipeline*

@@

Consider a commonly used drought index: Standardized Precipitation-Evapotranspiration Index (SPEI)  [@spei]. Its construction involves: 1) calculating potential evapotranspiration (PET) from average temperature, `tavg`, and its difference, `d`, with precipitation, `prcp`, 2) aggregating difference series with a time scale, 3) fitting the aggregated series with a chosen distribution to get density values, and 4) converting the density values to normal quantiles. Under the pipeline approach, SPEI will be constructed as:

```{r eval = FALSE, echo = TRUE}
DATA %>% 
  trans_pet(method = thornthwaite, col = tavg) %>% 
  aggregate(scale = 12, var = d) %>% 
  normalise(dist = gamma, fit_method = lmom) %>% 
  augment(var = .agg_prcp)
```

@@

*Change the examples to two scales values, compare how indices look like*

Here each command corresponds to one step described above with additional arguments specific to the step. 
The pipeline construction produces the same index value as command like `spei(...)`, as shown in @fig-toy-example, but with additional benefit:

@@

  *everyone is providing single command function, this is also the same for spi*

  *may not worthwhile if only one index, as long as start changing*

  *just want to calculate index, single function is fine*

```{r fig-toy-example, eval = TRUE, echo = FALSE}
#| fig-align: center
#| fig-cap: "Standardised Precipitation Evapotranspiration Index (SPEI) calculated with two distribution fits in Step 3 described above: generalised logistic (`glogist`) and generalised extreme value (`gev`) distribution, using the `wichita` data from the package `SPEI`."
#| out-height: 30%
#| out-width: 100%
knitr::include_graphics(here::here("figures/toy-example-spei.png"))
```

@@

@@

```{r}
#| fig-align: center
#| fig-cap: "Diagram of pipeline steps for index construction."
#| out-height: 90%
#| out-width: 90%
knitr::include_graphics(here::here("figures/pipeline-steps.png"))
```

### 2022-11-03 paper/paper.tex
change the small example to compare between two distributions

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmathfont[]{Latin Modern Math}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[normalem]{ulem}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, breakable, sharp corners, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

\emph{first sentence: tailor to their indices without thinking about the
big picture}

\sout{While every jumbled index jumbles in its own way,
{[}perspicuous{]} clearly-laied out indices are all alike.} This paper
proposes a data pipeline for index construction. By recognising the
common steps shared by many indices, we develop a pipeline that breaks
down index construction into multiple modules and allow operations in
various modules to be combined like building blocks to construct
indices. The pipeline approach is general while adaptable to most index
construction. It allows indices to be created, studied, and compared in
a structured tidy form and enables statistical analysis of indices to be
performed easily: More specifically, it enables researchers to 1)
validate the indices calculated from external organisations, 2) unify
various indices under the same framework for computing, 3) swap or
adjust individual steps in the index construction to study their
contribution, 4) calculate uncertainty on indices through bootstrap or
others, 5) enhance existing indices through comparing and studying their
statistical properties, and finally, 6) propose new indices from
combining different steps in existing indices.

@@

\hypertarget{sec-data-pipeline-in-r}{%
\section{Data pipeline in R}\label{sec-data-pipeline-in-r}}

@@

Consider a commonly used drought index: Standardised Precipitation Index
(SPI) (McKee, Doesken, and Kleist 1993). Its construction involves: 1)
aggregating monthly precipitation (\texttt{prcp}) with a time scale, 2)
fitting the aggregated series with a chosen distribution to get density
values, and 3) converting the density values to normal quantiles. Under
the pipeline approach, SPI will be constructed as:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{DATA }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{aggregate}\NormalTok{(}\AttributeTok{scale =} \DecValTok{12}\NormalTok{, }\AttributeTok{var =}\NormalTok{ prcp) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{normalise}\NormalTok{(}\AttributeTok{dist =}\NormalTok{ gamma, }\AttributeTok{var =}\NormalTok{ prcp, }\AttributeTok{fit\_method =}\NormalTok{ lmom) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{augment}\NormalTok{(}\AttributeTok{var =}\NormalTok{ .agg\_prcp)}
\end{Highlighting}
\end{Shaded}

Here each command corresponds to one step described above with
additional arguments specific to the step. The pipeline construction
produces the same index value as command like \texttt{spi(...)}, as
shown in Figure~\ref{fig-toy-example}, but with additional benefit:

@@

@@

{\centering \includegraphics[width=0.7\textwidth,height=0.7\textheight]{../figures/toy-example-spi.png}

@@

\caption{\label{fig-toy-example}Standardised Precipitation Index (SPI)
calculated using two approaches: the pipeline construction (y-axis) and
the function \texttt{spi()} from the \texttt{SPEI} package (x-axis),
using the \texttt{wichita} data from the package \texttt{SPEI}. All the
points align diagnally, indicating the same numerical result from both
approaches.}

@@

\emph{put a diagram here to summarise the pipeline}

@@

each step as a point rather than subsection

@@

\leavevmode\vadjust pre{\hypertarget{ref-spi}{}}%
McKee, Thomas B, Nolan J Doesken, and John Kleist. 1993. {``The
Relationship of Drought Frequency and Duration to Time Scales.''} In
\emph{Proceedings of the 8th Conference of Applied Climatology},
179--84. Anaheim, CA: American Meteorological Society.

@@

>>>>

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmathfont[]{Latin Modern Math}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, enhanced, boxrule=0pt, interior hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, sharp corners]}{\end{tcolorbox}}\fi

@@

This paper proposes a data pipeline for index construction. By
recognising the common steps shared by many indices, we develop a
pipeline that breaks down index construction into multiple modules and
allow operations in various modules to be combined like building blocks
to construct indices. The pipeline approach is general while adaptable
to most index construction. It allows indices to be created, studied,
and compared in a structured tidy form and enables statistical analysis
of indices to be performed easily: More specifically, it enables
researchers to 1) validate the indices calculated from external
organisations, 2) unify various indices under the same framework for
computing, 3) swap or adjust individual steps in the index construction
to study their contribution, 4) calculate uncertainty on indices through
bootstrap or others, 5) enhance existing indices through comparing and
studying their statistical properties, and finally, 6) propose new
indices from combining different steps in existing indices.

@@

\hypertarget{sec-data-pipeline-in-r}{%
\section{Data pipeline}\label{sec-data-pipeline-in-r}}

\emph{Think about if there is another word for data pipeline}

@@

Consider a commonly used drought index: Standardized
Precipitation-Evapotranspiration Index (SPEI) (Vicente-Serrano,
Beguería, and López-Moreno 2010). Its construction involves: 1)
calculating potential evapotranspiration (PET) from average temperature,
\texttt{tavg}, and its difference, \texttt{d}, with precipitation,
\texttt{prcp}, 2) aggregating difference series with a time scale, 3)
fitting the aggregated series with a chosen distribution to get density
values, and 4) converting the density values to normal quantiles. Under
the pipeline approach, SPEI will be constructed as:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{DATA }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{trans\_pet}\NormalTok{(}\AttributeTok{method =}\NormalTok{ thornthwaite, }\AttributeTok{col =}\NormalTok{ tavg) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{aggregate}\NormalTok{(}\AttributeTok{scale =} \DecValTok{12}\NormalTok{, }\AttributeTok{var =}\NormalTok{ d) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{normalise}\NormalTok{(}\AttributeTok{dist =}\NormalTok{ gamma, }\AttributeTok{fit\_method =}\NormalTok{ lmom) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{augment}\NormalTok{(}\AttributeTok{var =}\NormalTok{ .agg\_prcp)}
\end{Highlighting}
\end{Shaded}

\emph{Change the examples to two scales values, compare how indices look
like}

Here each command corresponds to one step described above with
additional arguments specific to the step. The pipeline construction
produces the same index value as command like \texttt{spei(...)}, as
shown in Figure~\ref{fig-toy-example}, but with additional benefit:

@@

\emph{everyone is providing single command function, this is also the
same for spi}

\emph{may not worthwhile if only one index, as long as start changing}

\emph{just want to calculate index, single function is fine}

@@

{\centering \includegraphics[width=1\textwidth,height=0.3\textheight]{../figures/toy-example-spei.png}

@@

\caption{\label{fig-toy-example}Standardised Precipitation
Evapotranspiration Index (SPEI) calculated with two distribution fits in
Step 3 described above: generalised logistic (\texttt{glogist}) and
generalised extreme value (\texttt{gev}) distribution , using the
\texttt{wichita} data from the package \texttt{SPEI}.}

@@

@@

{\centering \includegraphics[width=0.9\textwidth,height=0.9\textheight]{../figures/pipeline-steps.png}

\caption{Diagram of pipeline steps for index construction.}

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-spei}{}}%
Vicente-Serrano, Sergio M., Santiago Beguería, and Juan I. López-Moreno.
2010. {``A {Multiscalar} {Drought} {Index} {Sensitive} to {Global}
{Warming}: {The} {Standardized} {Precipitation} {Evapotranspiration}
{Index}.''} \emph{Journal of Climate} 23 (7): 1696--1718.
\url{https://journals.ametsoc.org/view/journals/clim/23/7/2009jcli2909.1.xml}.
